Stochastic Variance-Reduced Policy Gradient

### 2018-02-09 svrpg.tex
Corrected appendices experiments setting

\section{Conclusion}
In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
The control problem considered in the paper has a series of difficulties that are not common in SL.
Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy-dependent.
We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
}
\todomatout{Repetition of Section 6?}
Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.

@@

\begin{restatable}[]{lemma}{l-smooth-g}\label{lemma:gsmooth}
Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, $g(\tau\vert\vtheta)$ is Lipschitz continuous with Lipschitz constant $L_g$, \ie for any trajectory $\tau\in\Tspace$:
\[
	\norm[2]{g(\tau\vert\vtheta)-g(\tau\vert\vtheta')}^2 \leq L_g\norm[2]{\vtheta-\vtheta'}^2.
\]
\end{restatable}
\begin{proof}
        For both REINFORCE and G(PO)MDP, $g(\tau\vert\vtheta)$ is a linear combination of terms of the kind $\nabla \log \pi_{\vtheta}(a_t\vert s_t)\gamma^t r_t$ \cite{peters2008reinforcement}. These terms have bounded gradient from the second inequality of Assumption \ref{ass:bounded_score} and the fact that $|r_t|\leq R$. Bounded gradient implies Lipschitz continuity. Finally, the linear combination of Lipschitz continuous functions is Lipschitz continuous.
\end{proof}

\begin{restatable}[]{lemma}{bounded-g}\label{lemma:gbound}
Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, for every $\tau\in\Tspace$ and $\vtheta\in\Theta$, there is a positive constant $\Gamma<\infty$ such that:
\[
	\norm[2]{g(\tau\vert\vtheta)}^2 \leq \Gamma.
\]
\end{restatable}
\begin{proof}
For REINFORCE we have, from Assumption \ref{ass:bounded_score}:
\begin{align*}
	\norm[2]{g(\tau\vert\vtheta)}^2 &=
	\norm[2]{\score{\vtheta}{\tau}R(\tau)}^2 \\
	&=\norm[2]{\left(\sum_{t=0}^{H-1}\score{\vtheta}{a_t\vert s_t}\right)\left(\sum_{t=0}^{H-1}\gamma^t r_t\right)}^2\leq H^2G^2\frac{(1-\gamma^H)^2}{(1-\gamma)^2}R^2\dim(\vTheta) \coloneqq \Gamma
\end{align*}
For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\nabla \log \pi_{\vtheta}(a_t | s_t)\gamma^t r_t$ \cite{baxter2001infinite,peters2008reinforcement}, the above bound still holds.
\end{proof}

@@

\section{Experimental Details}\label{app:exp}
Setting specifications for all the used environments:
\begin{enumerate}
	\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, cart velocity $\dot{x}$ and pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body. Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$. The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
	\item \emph{Mujoco Swimmer}: 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles. 2-dimensional action space: the two momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$. The episodes terminates when the number of time steps T is greater than 500.
	\item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles. 6-dimensional action space: the 6 momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$. The episodes terminates when the number of time steps T is greater than 500.
\end{enumerate}

All the experiments' parameters are reported in the following table:

\centering
\begin{tabular}{| l | c  c  c |}
	\hline	
	& Cart-Pole & Swimmer & Half Cheetah \\
	\hline
	State space dimension  & 4 & 13 & 20 \\
	Control space dimension & 1 & 2 & 6 \\
	NN Hidden layers & 8 & 32x32 & 32x32 \\
	Learning rate SVRPO & $5\cdotp10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
	Learning rate GPOMDP & $10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
	Snapshot trajectories & 100 & 100 & 100 \\
	Sub-iterations trajectories & 10 & 10 & 10 \\
	Max sub-iterations & 50 & 20 & 20 \\
	Trajectories lengths& 500 & 500 & 500 \\
	Discount $(\gamma)$& 0.99 & 0.995 & 0.99 \\
	Total trajectories& 10000 & 20000 & 50000 \\
	\hline  
\end{tabular}

>>>>

\section{Conclusion}
In this paper, we introduced SVRPG, a variant of SVRG designed explicitly for RL problems.
The control problem considered in the paper has a series of difficulties that are not common in SL.
Among them, non-concavity and approximate estimates of the FG have been analysed independently in SL~\citep[\eg][]{allen2016variance,reddi2016stochastic,harikandeh2015stopwasting} but never combined.
Nevertheless, the main issue in RL is the non-stationarity of the sampling process since the distribution underlying the objective function is policy-dependent.
We have shown that by exploiting importance weighting techniques, it is possible to overcome this issue and preserve the unbiasedness of the corrected gradient.
We have additionally shown that, under mild assumptions that are often verified in RL applications, it is possible to derive convergence guarantees for SVRPG.
{\color{red}SVRG has been previously used in RL for policy evaluation~\citep{du2017svrgpe} and as alternative optimization method for TRPO~\citep{xu2017svrgtrpo}.
The first case is much more similar to the SL case since the problem is stationary and may be made concave, removing several issues we have faced here.
In the second case, it was used as a better optimization technique to solve the per-iteration problem imposed by TRPO.
}
\todomatout{Repetition of Section 6?}
Finally, we have empirically shown that practical variants of the theoretical SVRPG version can outperform classical actor-only approaches on benchmark tasks.
\todomat{add future work}

@@

\begin{restatable}[]{lemma}{l-smooth-g}\label{lemma:gsmooth}
Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, $g(\tau\vert\vtheta)$ is Lipschitz continuous with Lipschitz constant $L_g$, \ie for any trajectory $\tau\in\Tspace$:
\[
	\norm[2]{g(\tau\vert\vtheta)-g(\tau\vert\vtheta')}^2 \leq L_g\norm[2]{\vtheta-\vtheta'}^2.
\]
\end{restatable}
\begin{proof}
        For both REINFORCE and G(PO)MDP, $g(\tau\vert\vtheta)$ is a linear combination of terms of the kind $\nabla \log \pi_{\vtheta}(a_t\vert s_t)\gamma^t r_t$ \cite{peters2008reinforcement}. These terms have bounded gradient from the second inequality of Assumption \ref{ass:bounded_score} and the fact that $|r_t|\leq R$. If a baseline is used in REINFORCE or G(PO)MDP, we only need the additional assumption that said baseline is bounded.
        Bounded gradient implies Lipschitz continuity. Finally, the linear combination of Lipschitz continuous functions is Lipschitz continuous.
\end{proof}

\begin{restatable}[]{lemma}{bounded-g}\label{lemma:gbound}
Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, for every $\tau\in\Tspace$ and $\vtheta\in\Theta$, there is a positive constant $\Gamma<\infty$ such that:
\[
	\norm[2]{g(\tau\vert\vtheta)}^2 \leq \Gamma.
\]
\end{restatable}
\begin{proof}
For REINFORCE we have, from Assumption \ref{ass:bounded_score}:
\begin{align*}
	\norm[2]{g(\tau\vert\vtheta)}^2 &=
	\norm[2]{\score{\vtheta}{\tau}R(\tau)}^2 \\
	&=\norm[2]{\left(\sum_{t=0}^{H-1}\score{\vtheta}{a_t\vert s_t}\right)\left(\sum_{t=0}^{H-1}\gamma^t r_t\right)}^2\leq H^2G^2\frac{(1-\gamma^H)^2}{(1-\gamma)^2}R^2\dim(\vTheta) \coloneqq \Gamma
\end{align*}
For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\nabla \log \pi_{\vtheta}(a_t | s_t)\gamma^t r_t$ \cite{baxter2001infinite,peters2008reinforcement}, the above bound still holds.
If a baseline is used in REINFORCE or G(PO)MDP, we only need the additional assumption that said baseline is bounded.
\end{proof}

@@

\section{Experimental Details}\label{app:exp}
Used environments:
\begin{enumerate}
	\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, cart velocity $\dot{x}$ and pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body. Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$. The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
	\item \emph{Mujoco Swimmer}: 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles. 2-dimensional action space: the two momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2] a^2$. The episodes terminates when the number of time steps T is greater than 500.
	\item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles. 6-dimensional action space: the 6 momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 0.05\norm[2] a^2$. The episodes terminates when the number of time steps T is greater than 500.
\end{enumerate}

Refer to \cite{duan2016benchmarking}  for more details about GPOMDP in Half Cheetah task.
All the experiments' parameters are reported in the following table:

\centering
\begin{tabular}{| l | c  c  c |}
	\hline	
	& Cart-Pole & Swimmer & Half Cheetah \\
	\hline
	State space dimension  & 4 & 13 & 20 \\
	Control space dimension & 1 & 2 & 6 \\
	NN Hidden layers & 8 & 32x32 & 100x50x25 \\
	Adam learning rate SVRPO & $5\cdotp10^{-2}$ & $10^{-3}$ & $10^{-3}$ \\
	Adam learning rate GPOMDP & $10^{-2}$ & $10^{-3}$ & $10^{-2}$ \\
	Snapshot batch size & 100 & 100 & 100 \\
	Sub-iterations batch size & 10 & 10 & 10 \\
	GPOMDP batch size & 10 & 10 & 100 \\
	Max number of sub-iterations & 50 & 20 & 20 \\
	Episodes horizon& 100 & 500 & 500 \\
	Baseline& No & No & Yes \\
	Discount $(\gamma)$& 0.99 & 0.995 & 0.99 \\
	Total number of trajectories& 10000 & 20000 & 50000 \\
	\hline  
\end{tabular}
